---
title: "election results"
author: "Kuba"
date: "9/12/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
* **The packages to be used**
``` {r packages}
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(dplyr))
options(tibble.print_min=Inf); sixdec <- function(x) trimws(format(x, nsmall=6))
```

## function calculating the distribution of seats
(*the main function*)
```{r distr}
distr <- function(res) {
  seatsTot <- 20
  mis <- 0.5 ## shortlist threshold
  mip <- 0.05 ## party threshold
  nmin <- floor(Vt*mip)
  nmax <- ceiling(Vt*(1-mip))
  votes <- res[1:5]
  Vt <- res[6]
  
if(sum(votes) > Vt) {stop('voter turnout < total votes earned, ENTER OTHER DATA')}
  if(sum(votes)< mis*Vt) {warning('shortlist is less than 50%, reelection')}
  votes <- tibble(party=c("Communists","KPRF","LDPR","SR","ER"), votes=votes)
  for (i in 1:nrow(votes)) {if(votes$votes[i] >= nmax) {
    print(votes[i,])
    warning('there is a top-winner with more than 95%, reelection')}}
  calc <- tibble(party=rep(c("Communists","KPRF","LDPR","SR","ER"),(seatsTot -1)),
                 votes =rep(votes$votes, (seatsTot -1)), annex=0, denom=0, seats=0)
  calc[1:5,] <- calc[1:5,] %>%
    mutate(annex=if_else(votes %in% c(nmin:Vt), votes/2, 0), denom=2)
  
  for(i in 3:seatsTot) {
    calc[(nrow(votes)*(i-2)+1):(nrow(votes)*(i-1)),] <-
      calc[(nrow(votes)*(i-2)+1):(nrow(votes)*(i-1)),] %>%
      mutate(annex=if_else(votes %in% c(nmin:Vt), votes/i, 0), denom=i)
  }
  
  calc <- calc %>% filter(annex !=0) %>%
    arrange(desc(annex), desc(votes)) %>% mutate(annex=sixdec(annex))
  print('all denominators')
  print(calc)
  top <- calc %>% mutate(annex=as.numeric(annex)) %>%
    arrange(desc(annex), desc(votes)) %>%
    slice(1:seatsTot) %>% group_by(party) %>%
    mutate(party, seats = n()) %>% mutate(across(seats, as.numeric)) %>%
    ungroup() %>% arrange(desc(annex), desc(votes))
  print('top20 denominators & seats')
  print(top)
  
  byes <- c(which(!(unique(votes$party) %in% unique(calc$party))))
  distr <- top %>% select(party,seats) %>% distinct()
  if(length(byes)>0){
    distr <- distr %>% add_row(party=votes$party[byes], seats=0)     
  }
  
  losers <-c(which(!(unique(calc$party) %in% unique(top$party))))
  if(length(losers)>0) {
    distr <- right_join(distr, votes, by= c("party")) %>%
      arrange(desc(votes)) %>%
      add_row(party = "Vt", seats= sum(distr$seats, na.rm = TRUE), votes = Vt)
    print('provisional distribution')
    print(distr)
    
    for(i in 1:length(losers)){
      helpers <- top %>%
        mutate(seats = na_if(seats,1)) %>%
        arrange(annex,votes) %>%
        slice(1) %>% select(party)
      top <- top %>%
        mutate(seats=if_else(party %in% helpers, seats-1, seats))
    }
    
    losers <- unique(calc$party)[losers]
    losers <- calc %>% filter(party %in% losers) %>%
      mutate(annex=as.numeric(annex), seats=1) %>%
      group_by(party) %>% mutate(annex=max(annex)) %>%
      filter(row_number()==1)
    top <- top %>%
      mutate(seats=if_else(is.na(seats),1,seats))
    top <- bind_rows(top, losers) %>% arrange(desc(annex), desc(votes))
    warning('losers were here, calculated by paragraph ยง63.5 rule')
    print(c('losers:', losers$party))
    print(c('helpers:', helpers$party))
    distr <- top %>% select(party,seats) %>% distinct()
  }
  
  distr <- right_join(distr, votes, by= c("party")) %>%
    mutate(seats=if_else(is.na(seats),0,seats)) %>%
    arrange(desc(seats), desc(votes)) %>%
    add_row(party = "Vt", seats= sum(distr$seats, na.rm = TRUE), votes = Vt)
  print('final disrtibution')
  distr
}
```

===

> **example with a given protocol result**:

* **for the program to work**
  * set **the following data**: `Communists`, `KPRF`, `LDPR`, `SR`, `ER`,`Vt`
  (*votes earned by each party and voter turnourt*),
  * **generate the vector**: `res <- c(Communists,KPRF,LDPR,SR,ER,Vt)`,
  * **run the fuction**: `distr(res)`
```{r given}
## Communists: 2300, KPRF: 13420, LDPR: 7679, SR: 500 ,ER: 16322, turnout: 40221
suppressWarnings(rm(Communists, ER, KPRF,LDPR,res,SR,Vt))
Communists <- 2300
KPRF <- 13240
LDPR <- 7679
SR <- 500
ER <- 16322
Vt <- 40221 ## voter turnout
res <- c(Communists,KPRF,LDPR,SR,ER, Vt)
distr(res)
```

### function generating a random set
(*for experiments*)
```{r random}
## by default, Nv (number of voters) == 95043
rndres <- function(Nv=95043) {
        r <- runif(6)
        r <- sort(r[1:6])
        r[6] <- r[6] - r[5]
        r[5] <- (r[5]-r[4])
        r[4] <- (r[4]-r[3])
        r[3] <- (r[3]-r[2])
        r[2] <- (r[2]-r[1])
        Communists <- round(Nv*r[1])
        KPRF <- round(Nv*r[2])
        LDPR <- round(Nv*r[3])
        SR <- round(Nv*r[4])
        ER <- round(Nv*r[5])
        Vt <- round(sum(r)*Nv)
        rndres <- c(Communists,KPRF,LDPR,SR,ER, Vt)
        names(rndres) <- c("Communists","KPRF","LDPR","SR","ER", "Vt")
        print(rndres)
        rndres
}
```
===

> **random results example**:

* **run the `distr` function on random results**: `exper <- distr(rndres())`,
  * *OPTIONAL: set **number of voters (Nv)**: `exper <- distr(rndres(Nv))`
  (*by default Nv == 95043*)
  * **get the final result**: `exper`
```{r rex}
exper <- distr(rndres())
exper
```
